## Packages needed: tidyverse, sf, rnaturalearth, countrycode, ggrepel, rnaturalearthdata
```{r}
install.packages("tidyverse")
install.packages("sf")
install.packages("rnaturalearth")
install.packages("ggrepel")
install.packages("countrycode")
install.packages("ggplot2")
```

```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(ggrepel)
library(countrycode)
library(ggplot2)
library(scatterplot3d)
```

#Importing the datasets
```{r}
avgIQ <- read.csv("C:/Users/Leila/OneDrive/Desktop/Rproject/AvgIQpercountry.csv")
EduExp <- read.csv("C:/Users/Leila/OneDrive/Desktop/Rproject/expenditureoneducation.csv")
IQclass <- read.csv("C:/Users/Leila/OneDrive/Desktop/Rproject/IQclassification.csv")
```

# Display the datasets
```{r}
head(avgIQ)
head(IQclass)
head(EduExp)
```

# Remove both duplicates and n/a
```{r}
avgIQ <- avgIQ %>%
  distinct() %>%
  drop_na()
IQclass <- IQclass %>% 
  distinct() %>% 
  drop_na()
EduExp <- EduExp %>% 
  drop_na()

sum(is.na(avgIQ))
sum(duplicated(avgIQ))
sum(is.na(IQclass))
sum(duplicated(IQclass))
sum(is.na(EduExp))
sum(duplicated(EduExp))
```

# Count number of unique observations
```{r}
unique(avgIQ)
unique(IQclass)
unique(EduExp)
```

# Formatting the dataset: changing population into a numeric variable.
```{r}
avgIQ$Population...2023 <- as.numeric(as.character(avgIQ$Population...2023))
class(avgIQ$Population...2023)
```

# Manipulating the datasets: merging GovExp to avgIQ dataset
```{r}
vars <- c("id", "Rank", "Country", "Average.IQ", "Continent", "Literacy.Rate", "Nobel.Prices", "HDI..2021.", "Mean.years.of.schooling...2021", "GNI...2021", "Population...2023", "Latitude", "Longitude" )

avgIQ <- avgIQ %>% 
  left_join(EduExp, by = "id") %>% 
  select(all_of(vars), "GovExp")
```

#Removing what I don't need anymore
```{r}
remove(vars)
remove(EduExp)
```

# Data Transformation to create df for level of IQ
```{r}
IQdataset <- avgIQ %>% 
  group_by(id, Country) %>% 
  mutate(IQlevel = case_when(
    Average.IQ >= 178.00 ~ "Genius",
    Average.IQ >= 160.00 ~ "Almost genius",
    Average.IQ >= 140.00 ~ "Very gifted",
    Average.IQ >= 130.00 ~ "Gifted",
    Average.IQ >= 120.00 ~ "Moderately gifted",
    Average.IQ >= 120.00 ~ "Above average",
    Average.IQ >= 90.00 ~ "Average intelligence",
    Average.IQ >= 85.00 ~ "Below average",
    Average.IQ >= 70.00 ~ "Borderline intellectual functioning",
        Average.IQ >= 50.00 ~ "Mid intellectual disability",
    Average.IQ >= 35.00 ~ "Moderate intellectual disability",
        Average.IQ >= 20.00 ~ "Severe intellectual disability",
    Average.IQ >= 0.00 ~ "Profound intellectual disability", )) 

head(IQdataset)

```

```{r}
remove(avgIQ)
```

# Ordering the IQ dataset 
```{r}
IQdataset <- IQdataset %>% 
  relocate(Continent, Latitude, Longitude, Population...2023, .before = Average.IQ) %>% 
  relocate(Rank, IQlevel, .after = Average.IQ) %>% 
  relocate(HDI..2021., .before = GNI...2021)
head(IQdataset)
```

# I want to calculate the percentages I have in my sample in order to confirm its reliability.
```{r}
# First, I calculate frequencies:
ftable <- fct_count(IQdataset$IQlevel)
# then, I calculate the percentages:
round(prop.table(ftable$n)*100, digits = 2)
```


## WORLD MAP
#I need the world data 
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

# Selecting only the data that I want to plot in the map: country and iq level
```{r}
data <- IQdataset %>% 
  group_by(Country) %>% 
  summarize(Average_IQ = IQlevel) %>% 
  mutate(Iso3 = countrycode::countrycode(
    sourcevar = Country,
    origin = "country.name",
    destination = "iso3c")
  )
```

# Dropping Granade since I don't have it in the world dataframe
```{r}
data <- drop_na(data)
sum(is.na(data))
```

# Merging the world and the data dataframes trough ISO code
```{r}
IQmap_data <-world %>% 
  select(geometry, name, iso_a3) %>% 
  left_join(data, by = c("iso_a3" = "Iso3")) %>% 
  drop_na()
```

```{r}
remove(data)
```

# MAPPING
```{r}
world_map <- world %>% 
  filter(admin != "Antarctica") %>% 
  st_transform(crs = "+proj=robin") %>% 
  ggplot() +
  geom_sf(color = "grey") +
  geom_sf(data = IQmap_data, aes(fill = Average_IQ)) +
  scale_fill_brewer(type = "seq", palette = 4, direction = -1, aesthetics = "fill") +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(),
        legend.position = "right") +
  labs(title = "World Countries IQ",
       x = NULL, y = NULL,
       caption = "Average IQ levels per Country")

world_map
```

# For having different maps depending on the IQ level
```{r}
level_maps <- world %>% 
  filter(admin != "Antarctica") %>% 
  st_transform(crs = "+proj=robin") %>% 
  ggplot() +
  geom_sf(color = "grey") +
  geom_sf(data = IQmap_data, aes(fill = "Average.IQ")) +
  facet_wrap(~Average_IQ) +
  scale_fill_manual(values = "royalblue") +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(),
        legend.position = "none") +
  labs(title = "Levels of IQ",
       x = NULL, y = NULL,
       caption = "Average IQ levels per Country")

level_maps
```


# #####################################################
##STATISTICAL ANALYSIS

## 1st HYPOTHESIS: the IQ depends on the geographical zone.

# One-way ANOVA to check whether a correlation between geographical zone and intelligence exists.
```{r}
Geo_IQ <- aov(Average.IQ ~ Continent, data = IQdataset)
summary(Geo_IQ)
```

# Double check using also IQ levels
```{r}
chisq.test(IQdataset$Continent, IQdataset$IQlevel)
```

# Boxplot of Average IQ over continents
```{r}
IQdataset %>% 
  ggplot(aes(Continent, Average.IQ)) + 
  geom_boxplot(fill="lightblue", colour="blue")
```

# Correlation Continent - number of nobel prices won (to corroborate the racism hypothesis)
```{r}
Geo.nobel <- aov(Nobel.Prices ~ Continent, data = IQdataset)
summary(Geo.nobel)
#box plot to see the different averages
IQdataset %>% 
  ggplot(aes(Continent, Nobel.Prices)) + 
  geom_col(fill="lightblue", colour="blue")
```
#with colour=blue I can see that many european countries are on the same level as most asian ones, so probably it is not the continent of belonging itself that influences the level of IQ. What is that determines the differences among Countries in the same continent?
```{r}
lm.IQxcontinent.literacyrate <-lm(Average.IQ ~ Literacy.Rate:Continent,  data = IQdataset)
summary(lm.IQxcontinent.literacyrate)

#In effect, the literacy rate changes depending on which Continent we take into account:
lm.Litratexcontinent <- lm(Literacy.Rate ~ Continent, data = IQdataset)
summary(lm.Litratexcontinent)
```

```{r}
#Depending on the literacy rate, a Continent will have a higher or lower IQ average
lm.IQ.literacyrate <-lm(Average.IQ ~ Literacy.Rate, data = IQdataset)
summary(lm.IQ.literacyrate)

ggplot (data = IQdataset, aes (Literacy.Rate, Average.IQ)) + 
  geom_point () + 
  geom_smooth (method='lm')
```
# We found out that Literacy rate actually determines the IQ score. Maybe, it is not only literacy rate that matters in a society to determine the IQ.


## ###########################################################
## 2nd HYPOTHESIS: the IQ may depend on the quality of life.

#ANOVA and box plots to see whether different geographical zones have different qualities of life.
```{r}
Geo_HDI <- aov(HDI..2021. ~ Continent, data = IQdataset)
summary(Geo_HDI)

Geo_GNI <- aov(GNI...2021 ~ Continent, data = IQdataset)
summary(Geo_GNI)

#Population is not significant
Geo_Pop <- aov(Population...2023 ~ Continent, data = IQdataset)
summary(Geo_Pop)
lm(Population...2023 ~ Continent, data = IQdataset)
summary(lm)
```
#Both HDI and GNI scores depend on the continent. Statistically significant relation at the 1% level.

```{r}
#Visual comparison: GNI, HDI, IQ
IQdataset %>% 
  ggplot(aes(Continent, GNI...2021)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim=c(0, 85000))

IQdataset %>% 
  ggplot(aes(Continent, HDI..2021.)) + 
  geom_boxplot()

IQdataset %>% 
  ggplot(aes(Continent, Average.IQ)) + 
  geom_boxplot()
```
## Since quality of life depends on continents, I want to check whether we find a direct correlation between quality of life and intelligence.

# First, I perform other anova tests with the IQ level.
```{r}
HDI_IQ <- aov(HDI..2021. ~ IQlevel, data = IQdataset)
summary(HDI_IQ)

GNI_IQ <- aov(GNI...2021 ~ IQlevel, data = IQdataset)
summary(GNI_IQ)

IQdataset %>% 
  ggplot(aes(IQlevel, HDI..2021.)) + 
  geom_boxplot()

IQdataset %>% 
  ggplot(aes(Continent, GNI...2021)) + 
  geom_boxplot()
```
# Trying also with the average IQ and performing some regression analyses
```{r}
lm.IQxGNI <- lm(Average.IQ ~ GNI...2021, data = IQdataset)
summary(lm.IQxGNI)

lm.IQxHDI <- lm(Average.IQ ~ HDI..2021., data = IQdataset)
summary(lm.IQxHDI)

final.lm <- lm(Average.IQ ~ HDI..2021. + GNI...2021 + Population...2023, data = IQdataset )
summary(final.lm)

scatter3D <- scatterplot3d(cbind(x=IQdataset$HDI..2021., z=IQdataset$GNI...2021, y=IQdataset$Average.IQ), xlab="HDI",
                           ylab="GNI", zlab="Average IQ", type ="p",
                           color="blue", angle=45, pch=16)

```
# Here we find out that the effect of the Gross National Income is overcome when adding the Human Development Index, which is highly statisticlaly significant, and the population, whcih is significant at the 1% level.
# Endogeneity: Maybe the reason for this finding is that GNI and HDI are directly related.
```{r}
lm.HDIxGNI <- lm(HDI..2021. ~ GNI...2021, data = IQdataset)
summary(lm.HDIxGNI)
```

```{r}
linear.plot <- ggplot (data = IQdataset, aes (HDI..2021., Average.IQ)) + 
  geom_point () + 
  geom_smooth (method='lm')

linear.plot
```





